\subsection{Definitions}
Source case-patient. Secondary attack rate was calculated as the pro- portion of new cases among contact events × 100 (8) \cite{Dixon2015}.

Similar to algorithms described previously,4,9,41 a case was classified as \textit{primary} if it was the first case investigated in an independent outbreak chain, or had no history of contact with a symptomatic human within three weeks before rash onset. Primary cases were assumed to be caused by an animal source. \textit{Onset} is defined as the first day of rash presentation and is considered the first day of infectiousness. A \textit{chain of transmission} was defined as the primary case, and all subsequent cases stemming from this initial infection. A \textit{cluster} of cases may contain chains arising from multiple primary cases.29 A case in the same chain or cluster with an onset less than seven days after exposure to the primary case were classified as \textit{co-primary case}s. Any cases with an onset 7-34 days after reported contact with a symptomatic primary case was classified as a \textit{secondary case} of the first generation, and assumed to be caused by a human source. The \textit{generation} of secondary cases was determined by their location within the temporally classified chains of transmission.29
\textit{Secondary attack rates} are defined as the proportion of contacts exposed to an active case that developed the disease within the known incubation period. Attack rates can be compared to the \textit{effective reproductive number} ($R_{eff}$), which is defined as the average number of secondary cases caused by each individual case, with regard to population susceptibility.10,27 This is as opposed to the \textit{basic reproductive number} ($R_{0}$), which is the average number of secondary cases caused by each case in a completely susceptible population.29 \textit{Contacts} are defined as individuals coming into direct contact with a case, and \textit{household contacts} are those residing in the same household. Contacts were stratified into before and after contacts, as defined above. 